TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-07 paper/draft.tex
abundances

While elemental abundance measurements from high resolution spectra of F, G, and K stars have achieved extremely high precisions (down to errors in the range of $0.01-0.03$\,dex; \citealt{Nissen2018}) enabled by the improvement of atmospheric models (citations),
the determination of precise metallicities for M dwarfs has remained a long standing challenge, due to the formation of complex diatomic and triatomic molecular features at M dwarf temperature regimes, particularly TiO and VO which dominate the optical, H$_2$O and CO which dominate the infrared, and hydride molecules (FeH, CaH, CrH, MgH, etc.) which form at late spectral types \citep{Allard1997}.

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star, to developed precise empirical relations (as good as $\sim$0.07\,dex) between the adopted metallicities of the FGK primary (under the assumption kinematically similar pairs are co-evolutionary; \citealt{Bonfils:2005}), and strong, isolated lines in the optical spectra of the M dwarf (e.g. \citealt{Terrien:2012}, \citealt{Rojas-Ayala:2012}, \citealt{Newton:2014}).
However these metallicity calibrations are inaccesible for analysis with \apogee, as the narrow infrared wavelength coverage does not include the relevant calibration indices used by existing studies (such as Ca I or KI lines in the optical), and blending of multiple atomic and molecular species at high resolution makes it impractical to identify and deconvolve new individual features for analysis.

>>>>

While elemental abundance measurements from high resolution spectra of F, G, and K stars have achieved extremely high precisions (down to errors in the range of $0.01-0.03$\,dex; \citealt{Nissen2018}) enabled by improvements in atmospheric models towards realistic 3D local thermodynamic equilibrium (LTE) assumptions \citep{Asplund2005}, and differential abundance techniques using equivalent widths \citep{Bedell2014},
the determination of precise metallicities for M dwarfs has remained a long standing challenge due to the formation of complex diatomic and triatomic molecular bands at M dwarf temperature regimes, particularly TiO and VO which dominate the optical, H$_2$O and CO which dominate the infrared, and hydride bands (FeH, CaH, CrH, MgH, etc.) which form at late spectral types \citep{Allard1997}.
In particular, atmospheric models fail to reproduce spectral features due to incomplete line lists and opacities, and the presence of millions of weak, blended transitions and lack of continuum in a spectrum make it difficult to deconvolve individual features and extract line strengths from equivalent widths,
creating fundamental limitations in efforts towards exploiting the information content available out of high resolution datasets using traditional modeling and empirical methods.

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star, to developed precise empirical relations (as good as $\sim$0.07\,dex) between the adopted metallicities of the FGK primary (under the assumption kinematically associated wide binaries are co-evolutionary; \citealt{Bonfils:2005}), and strong, isolated lines in the optical spectra of the M dwarf (e.g. \citealt{Terrien:2012}, \citealt{Rojas-Ayala:2012}, \citealt{Newton:2014}).
However these metallicity calibrations are inaccesible for analysis with \apogee, as the narrow infrared wavelength coverage does not include the relevant calibration indices used by existing studies (such as Ca I or KI lines in the optical), and blending of multiple atomic and molecular species at high resolution makes it impractical to identify and deconvolve new individual features for analysis.
